Parallel Computing for Constrained Global Optimization Problems \thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}

### 2018-05-11 2018/Journal of Supercomputing/JoS_Paper.tex
Before final cheking

\begin{abstract}
This paper considers a parallel algorithm for solving multiextremal problems with non-convex constrains. Within the framework of the approach proposed here, the solution of the initial multidimensional problem is reduced solving a set of linked one-dimensional problems in parallel. The parallelization scheme does not require a single control process since the solution of any one-dimensional problem is a solution of the initial multidimensional one. The non-convex constraints are accounted for by a novel index scheme without utilizing the ideas of penalty functions. The efficiency of the parallel algorithm was confirmed experimentally by solving numerically of several hundred multidimensional multiextremal problems with non-convex constrains generated randomly.
\keywords{Global optimization \and Constrained problems \and Non-convex constraints \and Dimensionality reduction \and Parallel algorithms}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

Parallel algorithms utilizing heuristic approaches and based in one way or another on the idea of random search are at present quite popular (see \cite{RefFerreiro,RefZhu,Korosec,Guerrero}). However, because of their probabilistic nature, the algorithms of such type guarantee the convergence to the global optimum solution in the probabilistic sense only, which makes them less attractive than deterministic parallel global optimization algorithms \cite{Jones2001,Paulavicius2011,Evtushenko2013}. As a rule, efficient deterministic algorithms are intended to solve unconstrained optimization problems. The problems with constraints are solved using the penalty function method that has a number of disadvantages. In particular, this method cannot be applied to the problems with the \textit{partially defined functions} (i.e. when any constraint is violated, the values of all the other functions of the problem remain undefined).

@@

One of the possible ways to overcome this disadvantage consists in using the multiple mapping $Y^S(x)=\{ y^1 (x),...,y^S(x)\}$ instead of the single evolvent $y(x)$. To construct the set $Y^S(x)$, different approaches can be used. For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^i(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$. The set of Peano curves thus constructed allows one to obtain $y', y''$ from $D$ for any close multidimensional images, which differ only in one coordinate, close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$.

@@

In this study, we will use a GKLS generator of the functions of arbitrary dimensionality proposed in \cite{Gaviano2003} to produce constrained problems. The scheme for generating constrained global optimization problems is proposed in \cite{Gergel2017}. Previous investigations have confirmed experimentally that the index method is not inferior to well-known analogues. A comparison of the method with the well known DIRECT method \cite{Jones2001} in solving unconstrained optimization problems has been performed in \cite{Barkalov2015,Barkalov2016}. In the present study, an experimental investigation is carried out of the speedup, which is obtained by the use of the index method in combination with the two-level parallelization scheme from \cite{Sidorov2015}.

@@

\bibitem{Barkalov2016}
Barkalov, K., Gergel, V.: Parallel global optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016)

>>>>

\begin{abstract}
This paper considers a parallel algorithm for solving multiextremal problems with non-convex constrains. Within the framework of the approach proposed here, the solution of the initial multidimensional problem is reduced to solving a set of linked one-dimensional problems in parallel. The parallelization scheme does not require a single control process since the solution of any one-dimensional problem is a solution of the initial multidimensional one. The non-convex constraints are accounted for by a novel index scheme without utilizing the ideas of penalty functions. The efficiency of the parallel algorithm was confirmed experimentally by solving numerically of several hundred multidimensional multiextremal problems with non-convex constrains generated randomly.
\keywords{Global optimization \and Constrained problems \and Non-convex constraints \and Dimensionality reduction \and Parallel algorithms}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

Parallel algorithms utilizing heuristic approaches and based in one way or another on the idea of random search are at present quite popular (see \cite{RefFerreiro,RefZhu,Korosec,Guerrero}). However, because of their probabilistic nature, the algorithms of such type guarantee the convergence to the global optimum solution in the probabilistic sense only, which makes them less attractive than deterministic parallel global optimization algorithms \cite{Jones2001,Paulavicius2011,Evtushenko2013,Strongin2000}. As a rule, efficient deterministic algorithms are intended to solve unconstrained optimization problems. The problems with constraints are solved using the penalty function method that has a number of disadvantages. In particular, this method cannot be applied to the problems with the \textit{partially defined functions} (i.e. when any constraint is violated, the values of all the other functions of the problem remain undefined).

@@

One of the possible ways to overcome this disadvantage consists in using the multiple mapping $Y^S(x)=\{ y^1 (x),...,y^S(x)\}$ instead of the single evolvent $y(x)$. To construct the set $Y^S(x)$, different approaches can be used. For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^i(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$. The set of Peano curves thus constructed allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$ for any close multidimensional images $y', y''$ from $D$, which differ only in one coordinate.

@@

In this study, we will use a GKLS generator of the functions of arbitrary dimensionality proposed in \cite{Gaviano2003} to produce constrained problems. The scheme for generating constrained global optimization problems is proposed in \cite{Gergel2017}. Previous investigations have confirmed experimentally that the index method is not inferior to well-known analogues. A comparison of the method with the well known DIRECT method \cite{Jones2001} in solving unconstrained optimization problems has been performed in \cite{Barkalov2015}. In the present study, an experimental investigation is carried out of the speedup, which is obtained by the use of the index method in combination with the two-level parallelization scheme from \cite{Sidorov2015}.

@@

%\bibitem{Barkalov2016} Barkalov, K., Gergel, V.: Parallel global optimization on GPU. J. Glob. Optim. 66(1), 3--20 (2016)
